---
title: "Neural Networks"
author: "Amin Fesharaki"
date: "11/18/2021"
output:
  word_document: default
  html_document: default
  pdf_document: default
---
#Neural Networks - Chapter 9 Problems
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nnet)
library(NeuralNetTools)
library(caret)
library(plyr)
library(C50) 
library(rpart)
```

``` {r}
bank_train <- read.csv(file = "/Users/datascience/Desktop/ADS 502 Data Sets/Website Data Sets/bank_marketing_training.csv")
bank_test <- read.csv(file = "/Users/datascience/Desktop/ADS 502 Data Sets/Website Data Sets/bank_marketing_test.csv")

````

#24. Prepare the data set for neural network modeling,including standardizing the variables.

``` {r}

# Training Data Set

#Turn non-numerical variables into factors 
bank_train$job <- factor(bank_train$job)
bank_train$marital <- factor(bank_train$marital)
bank_train$default <- factor(bank_train$default)
bank_train$housing <- factor(bank_train$housing)
bank_train$education <- factor(bank_train$education)
bank_train$loan <- factor(bank_train$loan)
bank_train$contact <- factor(bank_train$contact)
bank_train$month <- factor(bank_train$month)
bank_train$previous_outcome <- factor(bank_train$previous_outcome)
bank_train$response <- factor(bank_train$response)
# Change Response to yes and No
bank_train$response <- revalue (bank_train$response, c("yes"=1))
bank_train$response <- revalue (bank_train$response, c("no"=0))

# min - max Standardization 
bank_train$age.mm <- (bank_train$age - min(bank_train$age)) / (max(bank_train$age)- min(bank_train$age))

bank_train$duration.mm <- (bank_train$duration - min(bank_train$duration)) / (max(bank_train$duration)- min(bank_train$duration))

bank_train$campaign.mm <- (bank_train$campaign - min(bank_train$campaign)) / (max(bank_train$campaign)- min(bank_train$campaign))

bank_train$days_since_previous.mm <- (bank_train$days_since_previous - min(bank_train$days_since_previous)) / (max(bank_train$days_since_previous)- min(bank_train$days_since_previous))

bank_train$previous.mm <- (bank_train$previous - min(bank_train$previous)) / (max(bank_train$previous)- min(bank_train$previous))


bank_train$emp.var.rate.mm <- (bank_train$emp.var.rate - min(bank_train$emp.var.rate)) / (max(bank_train$emp.var.rate)- min(bank_train$emp.var.rate))

bank_train$cons.price.idx.mm <- (bank_train$cons.price.idx - min(bank_train$cons.price.idx)) / (max(bank_train$cons.price.idx)- min(bank_train$cons.price.idx))
                               
bank_train$cons.price.idx.mm <- (bank_train$cons.price.idx - min(bank_train$cons.price.idx)) / (max(bank_train$cons.price.idx)- min(bank_train$cons.price.idx)) 

bank_train$cons.conf.idx.mm <- (bank_train$cons.conf.idx - min(bank_train$cons.conf.idx)) / (max(bank_train$cons.conf.idx)- min(bank_train$cons.conf.idx)) 

bank_train$nr.employed.mm <- (bank_train$nr.employed - min(bank_train$nr.employed )) / (max(bank_train$nr.employed )- min(bank_train$nr.employed))                       
````

#25. Using the training dataset,create a neural network model to predict a customer’s Response using whichever predictors you think appropriate. Obtain the predicted responses.

``` {r}
# Creating Neural
nnet01 <- nnet (response ~ age.mm + job + marital + education + contact + housing + loan + duration.mm + previous.mm + previous_outcome, data = bank_train, size = 1, maxit = 100)
````
#26. Plot the neural network.

``` {r}
plotnet(nnet01)
# make predictions (returns probabilities)
bank_train$pred_prob <- predict(object = nnet01, newdata = bank_train)

````
#27. Evaluate the neural network model using the test dataset.Construct a contingency table to compare the actual and predicted values of Response.
``` {r}
# Test Data Set
#Turn non-numerical variables into factors 
bank_test$job <- factor(bank_test$job)
bank_test$marital <- factor(bank_test$marital)
bank_test$default <- factor(bank_test$default)
bank_test$housing <- factor(bank_test$housing)
bank_test$education <- factor(bank_test$education)
bank_test$loan <- factor(bank_test$loan)
bank_test$contact <- factor(bank_test$contact)
bank_test$month <- factor(bank_test$month)
bank_test$previous_outcome <- factor(bank_test$previous_outcome)
bank_test$response <- factor(bank_test$response)
# Change Response to yes and No
bank_test$response <- revalue (bank_test$response, c("yes"=1))
bank_test$response <- revalue (bank_test$response, c("no"=0))

# min - max Standardization 
bank_test$age.mm <- (bank_test$age - min(bank_test$age)) / (max(bank_test$age)- min(bank_test$age))

bank_test$duration.mm <- (bank_test$duration - min(bank_test$duration)) / (max(bank_test$duration)- min(bank_test$duration))

bank_test$campaign.mm <- (bank_test$campaign - min(bank_test$campaign)) / (max(bank_test$campaign)- min(bank_test$campaign))

bank_test$days_since_previous.mm <- (bank_test$days_since_previous - min(bank_test$days_since_previous)) / (max(bank_test$days_since_previous)- min(bank_test$days_since_previous))

bank_test$previous.mm <- (bank_test$previous - min(bank_test$previous)) / (max(bank_test$previous)- min(bank_test$previous))


bank_test$emp.var.rate.mm <- (bank_test$emp.var.rate - min(bank_test$emp.var.rate)) / (max(bank_test$emp.var.rate)- min(bank_test$emp.var.rate))

bank_test$cons.price.idx.mm <- (bank_test$cons.price.idx - min(bank_test$cons.price.idx)) / (max(bank_test$cons.price.idx)- min(bank_test$cons.price.idx))
                               
bank_test$cons.price.idx.mm <- (bank_test$cons.price.idx - min(bank_test$cons.price.idx)) / (max(bank_test$cons.price.idx)- min(bank_test$cons.price.idx)) 

bank_test$cons.conf.idx.mm <- (bank_test$cons.conf.idx - min(bank_test$cons.conf.idx)) / (max(bank_test$cons.conf.idx)- min(bank_test$cons.conf.idx)) 


                                 
bank_test$nr.employed.mm <- (bank_test$nr.employed - min(bank_test$nr.employed )) / (max(bank_test$nr.employed )- min(bank_test$nr.employed)) 


# make predictions (returns probabilities)
bank_test$pred_prob_test <- predict(object = nnet01, newdata = bank_test)
# convert to classes
bank_test$pred_test <- (bank_test$pred_prob_test > 0.5)*1

# performance metrics / Confusion Matrix
bank_test[c('response', 'pred_test')] <- lapply(bank_test[c('response', 'pred_test')], as.factor)
confusionMatrix(bank_test$pred_test, bank_test$response, positive='1')
cm


````

```{r}
 #Contingency Table
c.pred <- table(bank_test$response, bank_test$pred_test)
rownames(c.pred) <- c("Actual: No", "Actual: Yes")
colnames(c.pred) <- c("Predicted: No", "Predicted: Yes")
addmargins(A = c.pred, FUN = list(Total=sum), quiet = TRUE)

TN0 <- c.pred[1,1]
FN0 <- c.pred[2,1]
FP0 <- c.pred[1,2]
TP0 <- c.pred[2,2]

```

# 28. Which baseline model do we compare your neural network model against? Did it out- perform the baseline according to accuracy?
``` {r}
#????????????????????????#
````

#29. Using the same predictors you used for your neural network model, build models to predict Response using the following algorithms: (a) Cart (b) C5.0 (C) Naives Bayes

``` {r}

# Setting up Predictions
X = data.frame(age.mm = bank_test$age.mm, job = bank_test$job, marital = bank_test$marital, education = bank_test$education, contact = bank_test$contact, housing = bank_test$housing, loan =  bank_test$loan, duration.mm  = bank_test$duration.mm, previous.mm =  bank_test$previous.mm,  previous_outcome  = bank_test$previous_outcome)

````
#(a) Cart
``` {r}
# Cart Model trained by training data set
library(rpart)
cart  <- rpart(formula = response ~ age.mm + job + marital + education + contact + housing + loan + duration.mm + previous.mm + previous_outcome, data = bank_train, method = "class")

bank_test$pred_cart <- predict(object = cart, newdata = X)

# Predictions Test Data Set
Pred_cart = predict(object = cart, newdata = X, type = "class")
head(Pred_cart)
````
```{r}
# Evaluation Metrics for Cart

cart.pred <- table(bank_test$response, Pred_cart)
rownames(cart.pred) <- c("Actual: No", "Actual: Yes")
colnames(cart.pred) <- c("Predicted: No", "Predicted: Yes")
addmargins(A = cart.pred, FUN = list(Total=sum), quiet = TRUE)

```
```{r}
# Assigning General Form of Table to matrix values for Cart
TN1 <- cart.pred[1,1]
FN1 <- cart.pred[2,1]
FP1 <- cart.pred[1,2]
TP1 <- cart.pred[2,2]
```

# (b) C5.0
```{r}
# C5 model trained by training data set
c5 <- C5.0(response ~ age.mm + job + marital + education + contact + housing + loan + duration.mm + previous.mm + previous_outcome, data = bank_train)

# Predictions Test Data Set
Pred_c5 = predict(object = c5, newdata = X)
head(Pred_c5)
```


```{r}
# Evaluation Metrics for C5.0
c5.pred <- table(bank_test$response, Pred_c5)
rownames(c5.pred) <- c("Actual: No", "Actual: Yes")
colnames(c5.pred) <- c("Predicted: No", "Predicted: Yes")
addmargins(A = c5.pred, FUN = list(Total=sum), quiet = TRUE)

```

```{r}
# Assigning General Form of Table to matrix values for C5.0
TN2 <- c5.pred[1,1]
FN2 <- c5.pred[2,1]
FP2 <- c5.pred[1,2]
TP2 <- c5.pred[2,2]
```

# (C) Naives Bayes
```{r}
# Naives Bayes model trained by training data set
library(e1071)
nb01 <- naiveBayes(formula = response ~ age.mm + job + marital + education + contact + housing + loan + duration.mm + previous.mm + previous_outcome, data = bank_train)

# Predictions Test data set
Pred_NB <- predict(object = nb01, newdata = X)
head(Pred_NB)
```

```{r}
# Evaluation Metrics for Naives Bayes

nb.pred <- table(bank_test$response, Pred_NB)
rownames(nb.pred) <- c("Actual: No", "Actual: Yes")
colnames(nb.pred) <- c("Predicted: No", "Predicted: Yes")
addmargins(A = nb.pred, FUN = list(Total=sum), quiet = TRUE)
```

```{r}
# Assigning General Form of Table to matrix values for NB
TN3 <- nb.pred[1,1]
FN3 <- nb.pred[2,1]
FP3 <- nb.pred[1,2]
TP3 <- nb.pred[2,2]
```


#30. Compare the results of your neural network model with the three models from the previous exercise, according to the following criteria. Discuss in detail which model performed best and worst according to each criterion. (A) Accuracy (B) Sensitivity (C) Specificity 

```{r}
# Neural Network 
Acc0 <- (TN0 + TP0) / (TN0 + FN0 + FP0 + TP0) # Accuracy
Sens0 <- (TP0) / (FN0 + TP0) #Sensitivity
Spec0 <- (TN0) / (TN0 + FP0) # Specificity

# Cart Model
Acc1 <- (TN1 + TP1) / (TN1 + FN1 + FP1 + TP1) # Accuracy
Sens1 <- (TP1) / (FN1 + TP1) #Sensitivity
Spec1 <- (TN1) / (TN1 + FP1) # Specificity

# C5.0 Model
Acc2 <- (TN2 + TP2) / (TN2 + FN2 + FP2 + TP2) # Accuracy
Sens2 <- (TP2) / (FN2 + TP2) #Sensitivity
Spec2 <- (TN2) / (TN2 + FP2) # Specificity

# Naives Bayes
Acc3 <- (TN3 + TP3) / (TN3 + FN3 + FP3 + TP3) # Accuracy
Sens3 <- (TP3) / (FN3 + TP3) #Sensitivity
Spec3 <- (TN3) / (TN3 + FP3) # Specificity

cat ("---Neural Network----", "\nAccuracy = ", Acc0, "\nSensitivity = ", Sens0, "\nSpecificty=", Spec0)
cat ("\n---Cart Model---", "\nAccuracy = ", Acc1, "\nSensitivity = ", Sens1, "\nSpecificty=", Spec1)
cat ("\n---C5.0 Model---", "\nAccuracy = ", Acc2, "\nSensitivity = ", Sens2, "\nSpecificty=", Spec2)
cat ("\n---Naives Bayes---", "\nAccuracy = ", Acc3, "\nSensitivity = ", Sens3, "\nSpecificty=", Spec3)
```